---
title: "Monte Carlo simulation to approximate characteristics of Gamma or Unkonw distribution. "
author: "Zhaojie"
date: "8/19/2024"
output: word_document
---
```{r}


```{r,第一个例子只有单参数}
#设置参数
set.seed(32) # Initializes the random number generator so we can replicate these results. To get different random numbers, change the seed.
m = 100
a = 2.0
b = 1.0 / 3.0

```

```{r}
#从Gamma分布中取theta样本
theta = rgamma(n=m, shape=a, rate=b)
print(theta)
head(theta)
tail(theta)
```

```{r}
#给抽样的样本们绘直方图对比Gamma曲线
hist(theta, freq=FALSE)#frequency = false option so that it gives us a probability density. Rather than counts.
curve(dgamma(x=x, shape=a, rate=b), col="blue", add=TRUE)
```

```{r}
#用抽样theta来近似估计Expect Value E[theta]
#这就是Monte Carlo simulation
sum(theta)/m
mean(theta)#5.752
#该Gamma（2，1/3）分布的真实E[theta]=a/b  α/β
a/b #6   
#5.7与6接近
```

```{r}
#增加抽样数，样本量来提升Monte Carlo的预测力
set.seed(32) 
m = 10000
a = 2.0
b = 1.0 / 3.0
theta = rgamma(n=m, shape=a, rate=b)
print(theta)
head(theta)
tail(theta)
hist(theta, freq=FALSE)
curve(dgamma(x=x, shape=a, rate=b), col="red", add=TRUE)
sum(theta)/m
mean(theta)#6.02
a/b#6   6.02与6非常接近了
```

```{r}
#使用Monte Carlo Simulation估计Var(theta)
var(theta) #18.01
#真实Var(theta)值：a/b^2, α/β^2
a/b^2 #18  18.01与18非常接近

```

```{r}
#if we want to approximate the probability that theta is less than 5
#We can simply count how many of our simulated thetas are less than 5

ind = theta < 5.0 # set of indicators, TRUE if theta_i < 5
head(ind)
head(theta)
mean(ind) # automatically converts FALSE/TRUE to 0/1
#0.4974
```

```{r, pgamma(  )}
#Ture Gamma(2,1/3) probability that θ<5
pgamma(q=5.0, shape=a, rate=b)
#0.4963
#可见对于“probability that θ<5”的预测，0.4974与0.4963非常接近
```

```{r, qgamma(  ) & quantile function}
#What is the 0.9 quantile (90th percentile) of θ ？
#We can use the quantile function which will order the samples for us #and find the appropriate sample quantile.
quantile(theta,prob=0.9)
#90% 0.9 quantile (90th percentile)
#11.74426

#Ture value of 90percentile 0.9quantile
qgamma(p=0.9, shape=a, rate=b)
#11.66916
#可见对于“90percentile 0.9quantile”的预测，11.74与11.66非常接近


###反思这里原分布Beta(α，β)我们可以轻松算出真实总体参数
###但如果某分布epsilon(a,b,c)我们根本无法计算出参数，则必须用MC
```

```{r}
###Monte Carlo error###
#We can use the CLT to approximate how accurate our Monte Carlo estimates are. 
#For example, if we seek E(θ) then the sample mean θ¯∗
#approximately follows a normal distribution with mean E(θ) and #variance Var(θ)/m
#We will use the sample standard deviation divided by the square root #of m to approximate the Monte Carlo standard deviation.
```

```{r}
#error of MC approximate E[theta]
set.seed(32) 
m = 10000
a = 2.0
b = 1.0 / 3.0
theta = rgamma(n=m, shape=a, rate=b)
#calulate standard error
se = sd(theta) / sqrt(m)
2.0 * se # we are reasonably confident that (95%) the Monte Carlo estimate is no more than this far from the truth
mean(theta) - se
mean(theta) + se
#So 95% ceridale interval of estimate E[theta] is (5.979,6.064) 
```

```{r}
#error of approximate the "probability that theta is less than 5"
ind = theta < 5.0 # set of indicators, TRUE if theta_i < 5
head(ind)
head(theta)
mean(ind) # automatically converts FALSE/TRUE to 0/1
#Standard error of "probability that …" estimate
se=sd(ind)/sqrt(m)
2.0 * se # we are reasonably confident that (95%) the Monte Carlo estimate is no more than this far from the truth
mean(ind)-se
mean(ind)+se
#So 95% ceridale interval of estimate "Prob that theta >/< k" is (0.492,0.502) 
```




```{r，第二个例子，双参数，hierarchy model'}
1, Simulate Φi from Beta(2,2)
2, Simulate Yi from Binomial(10,Φi)
```

```{r}
m = 10e4 #10000 1万次

y = numeric(m) # create the vectors we will fill in with simulations
phi = numeric(m)
#Build a Loop
for (i in 1:m) {
  phi[i] = rbeta(n=1, shape1=2.0, shape2=2.0)
  y[i] = rbinom(n=1, size=10, prob=phi[i])
}

# which is equivalent to the following 'vectorized' code 运行更快
phi = rbeta(n=m, shape1=2.0, shape2=2.0)
y = rbinom(n=m, size=10, prob=phi)
```

```{r}
#If we are interested only in the marginal distribution of Y
#we can just ignore the draws for ϕ Beta(2,2)
#Treat the draws of Y as a sample from its marginal distribution.
#Here Y no longer binomial, actually it be beta-binomial

#Remember Conditional on phi Φ, Y follows a binomial distribution. 
#But unconditionally, the marginal distribution of y is not binomial. 
```

```{r}
table(y)    #Y是Beta-Binomial n=10 取10次 给K=1-10次列个表格
table(y)/m  #除以样本数，可以得到大致的每个k次成功的概率(占比)
mean(y) #均值在4.99
#注意，这是忽略了参数PhiΦ的Marginal Distribution of Y solely
```

```{r}
#给Marginal Distribution Y (Beta-Binomial)画图
plot(table(y)/m)
#更精细的画图
plot(prop.table(table(y)), ylab="P(y)", main="Marginal distribution of y")
```

